---
title: "Project 2"
author: "Honorio Vega, Andrew Sanchez, Bret Stine "
date: "April 24, 2018"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

```{r }
dat = read.csv("data/Video_Game_Sales_as_of_Jan_2017.csv")
dat$dtime = strptime(as.character(dat$Crash.Date.Time), "%m/%d/%Y %I:%M:%S %p", tz="America/New_York")
```

### Test changes
* this a test for push

###BRET WAS HERE --------> ****

### Intro: 
* We look at causes of crashes, how often accidents happen, examined drunk driving, miraculous survivors, locations of accidents, weather conditions, and traffic controls.

### Data: 
* We downloaded the Crash Reporting - Drivers Data in csv format from https://data.montgomerycountymd.gov/Public-Safety/Crash-Reporting-Drivers-Data/mmzv-x632 on March 13, 2018. 

### Initial data exploration:
* Upon initial exploration of the data, there are 32 columns and 65841 rows. All columns contain data of type 'Factor', except for longitude and latitude which are 'num'. Then we added another column, dat$dtime, with type POSIXlt to hold timestamps of the reported crash.

### Data cleaning and preprocessing:
* After looking over the data, we decided there wasn't any corrupted or NA data. There was a good amount of missing data for a few columns but we decided they were not significant enough to affect our work. We did do some preprocessing however, by adding a new dat$dtime column to hold POSIXlt type timestamps.

### Data exploration and visualization:
* See Questions 1-10 below.

### Conclusions:
* We learned how to use POSIXlt and work with timestamps.
* Data can easily be omitted to skew a narrative. For example, the most common statistic for distracted driving might be N/A or Not Distracted, but data scientists can choose to focus only on the specified reasons for distracted driving. 

