---
title: "Project 2 - Video Game Sales Prediction"
author: "Honorio Vega, Andrew Sanchez, Bret Stine"
date: "May 3, 2018"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

```{r }
<<<<<<< HEAD:output.Rmd
dat = read.csv("data/Video_Game_Sales_as_of_Jan_2017.csv")
```


=======
setwd('..')
dat = read.csv(paste0(getwd(),"/data/Video_Game_Sales_as_of_Jan_2017.csv"))
```

>>>>>>> c0860ba39ba05d8c6b017ca3a8a13cc20be749c6:templates/output.Rmd
### Intro: 
* We look at video games popularity, based on the ratings and reviews we will try to predict the sales of a game.

### Data: 
* We downloaded the Video Game data in csv format from https://www.kaggle.com/kendallgillies/video-game-sales-and-ratings/data on May 3, 2018. 

### Initial data exploration:
* Upon initial exploration of the data, there are 32 columns and 65841 rows. All columns contain data of type 'Factor', except for longitude and latitude which are 'num'. Then we added another column, dat$dtime, with type POSIXlt to hold timestamps of the reported crash.

### Data cleaning and preprocessing:
* After looking over the data, we decided there wasn't any corrupted or NA data. There was a good amount of missing data for a few columns but we decided they were not significant enough to affect our work. We did do some preprocessing however, by adding a new dat$dtime column to hold POSIXlt type timestamps.

### Data exploration and visualization:
* See Questions 1-10 below.

### Conclusions:
* We learned how to use POSIXlt and work with timestamps.
* Data can easily be omitted to skew a narrative. For example, the most common statistic for distracted driving might be N/A or Not Distracted, but data scientists can choose to focus only on the specified reasons for distracted driving. 

