---
title: "Project 2 - Video Game Sales Prediction"
author: "Honorio Vega, Andrew Sanchez, Bret Stine"
date: "May 3, 2018"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

### Goal
Our team's goal is to predict global video game sales based off of collected data in the past 15 years. 

### Intro(Need More)
The dataset comes from Kaggle.com and is a compilation of video games sales along with ratings. 

### Read & Preprocessing
We read the data in and begin preprocessing. First thing we do is get the total number of NAs values and see that there is 37396 rows of missing data. In order to keep our data recent and relevent we decided to use complete cases. Since many of the games with a year of release lower than 2000s tended to have missing data such as global sales, critic score, and user score.
```{r }
setwd('..')
dat = read.csv(paste0(getwd(),"/data/Video_Game_Sales_as_of_Jan_2017.csv"))
names(dat) = tolower(names(dat))
```
The data set contains 17416 rows of data with a majority of games being released between 2000s and above.

```{r}
set.seed(123)
before = nrow(dat)
before

par(mar=c(3,7,1,1))
barplot(table(dat$year_of_release), horiz=T, col="RED", las=1)
```

After using complete cases we retain 41% of our data that is recent and relevant to our prediction of global sales. 
```{r}
dat = dat[complete.cases(dat),]
after = nrow(dat)
after/before
```




